### 2020-11-18 latex/acrobot.tex
First writeup of acrobot energy gain proof for oscillations.

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
    Let 
    \[
        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
        ,
    \]
    be the energy of the simple pendulum obtained by setting \(I = 0\).
\begin{enumerate}
    \item There always exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
    \[
        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
            \mid E(q_u,p_u) < E(\pi,0) \right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\mu_0)d\sigma > 0\) for all 
    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \Omega := \left\{(q_u,p_u) \in \Gamma 
            \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
\end{enumerate}
\end{thm}
%\footnotetext{In fact, we can be more precise than this. 
%    Letting \(\epsilon > 0\), we need 
%    \(I \leq \min\left\{I_1(\epsilon),I_2\}\) where
%    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
%    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
%    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
%    \right)^{-1/2}\)
%    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
%    See Apendix \ref{apx:I-bounds} for details.
%}
The region \(\mathcal{O}\) is the oscillation domain of the simple pendulum
obtained by setting \(q_a = 0\);
\(\Omega\) includes \(\mathcal{O}\), alongside all rotations of this 
pendulum with momentum less than \(\bar{\mu}\). 
Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
that the acrobot will gain enough energy to begin performing giant motion;
the second result guarantees can achieve giants with momentum
\(|p_u| < \bar{\mu}\).

@@

\section{Proving the Acrobot Gains Energy}\label{sec:acrobot-proof}
% TODO: Prove this in sections
% TODO: You don't have to use the proof of theta_dot being positive. Stick to
% Manfredi's proof, and when you say "for I small enough" put in a footnote that
% says "in fact, we can be more precise and set bounds for I given as I <
% min{...}. See appendix for more detail."
% Then, we can have an appendix where we derive the bounds on I so that
% theta_dot is always positive. This will help the reader keep track and not
% bother them too much.
Proving Theorem \ref{thm:acrobot-energy-stabilization} is not a simple task.
In an effort to make the proof as clear as possible, we will break it down into
the following segments:
\begin{enumerate}
    \item Background on perturbation theory and averaging.
    \item Perturbation analysis for oscillations.
    \item Perturbation analysis for rotations.
\end{enumerate}

@@

\citet{khalil_nonlinear} considers a system of the form
\begin{equation}\label{eqn:khalil-setup}
    \begin{cases}
        \dot{x} = f(t,x,\epsilon), \\
        x(t_0) = \eta(\epsilon),
    \end{cases}
\end{equation}
where 
\(f : [t_0,t_1] \times D \times [-\epsilon_0,\epsilon_0] \rightarrow \R^n\) is
sufficiently smooth.
In the context of the acrobot, \(x\) is the set of coordinates which make
our analysis convenient;
\(\epsilon\) is our control parameter \(I\);
the function \(f\) is the constrained dynamics in \(x\)-coordinates;
and \(\eta(\epsilon) \equiv \eta_0\) is a constant initial condition.
To keep a notational reminder that we are applying perturbation theory to
the acrobot, we will henceforth use \(I\) in place of \(\epsilon\).

Setting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use this to compute the solution \(x_1(t,\eta_0)\) of
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}
Finally, one can compute the first-order Taylor approximation
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) is \(O(I^2)\), \ie,
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    .
\]
We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}\label{thm:khalil-perturbation}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem ~\eqref{eqn:khalil-perturbation-nominal} has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(I,\, |I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
As we have seen, the acrobot becomes a nominal pendulum that oscillates whenever the
nominal energy ~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\).
Hence, the domain of oscillations for the nominal pendulum is given by
\[
    \mathcal{O} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(\pi,0)\right\}
    ,
\]
where \(\Gamma \simeq \Sone\times\R\) is the constraint manifold of the
acrobot.
We wish to find a transformation \(T(q_u,p_u)\) on \(\mathcal{O}\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \Sone\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}) 
and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), we set \(\mu \in ]0,\pi[\).
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
The dynamics of the nominal pendulum arise from evaluating
\(f_\mu\) and \(f_\alpha\) at \(I = 0\).
MATLAB's symbolic toolbox evaluates the nominal dynamics as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
            {\mu_n^2 \sin(\alpha_n)^2}}
    .
\end{align}
Note that we write the nominal coordinates as \((\alpha_n,\mu_n)\) to
distinguish them from the acrobot's coordinates.
One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\).
This is expected of an oscillating single pendulum.
By continuity of ~\eqref{eqn:acrobot-mu-dot}, 
\(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for \(I\) small enough.

@@

%Using perturbation theory, we will find a first-order
%approximation of \(\hat{\mu}\). 
%We will show the dynamics of this approximation are \(\pi\)-periodic in
%\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
%decreases along a half-orbit \(\alpha \in [0,\pi]\).
%To complete the proof of energy gain for oscillations, we will demonstrate that
%\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
%This will imply that the the radius of orbits increases
%until all orbits within \(\mathcal{O}\) eventually leave \(\mathcal{O}\) in finite time.

In the spirit of perturbation analysis, we expand the solution of
~\eqref{eqn:muhat-dot} into
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha,\mu_0,I) = \hat{\mu}_0(\alpha,\mu_0) + I
    \hat{\mu}_1(\alpha,\mu_0)
    + R(\alpha,\mu_0,I)
    ,
\end{equation}
where \(R(\alpha,\mu_0,I)\) is \(O(I^2)\).
From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    ,
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

To prove this intuitive argument rigorously, we define the Poincar\'{e} section
\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

@@

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive,
monotonically increasing, and 
\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).

Suppose we initialize the acrobot at \((q_0,p_0) \in \mathcal{O}\).
Since on \(\mathcal{O}\) we have \(\dot{\alpha} > 0\), the orbit of the acrobot
in \(\Gamma\) spirals clockwise around the origin; 
if the spiral does not exit \(\mathcal{O}\), it must intersect the \(q_u\) axis.
Once it does, the Poincar\'{e} section \(P_\mathcal{O}(\cdot)\) tells us 
the orbit must spiral away from the origin. 
This implies there cannot be any closed orbits in \(\mathcal{O}\) and that the
origin is a repeller.

Once the orbit reaches an intersection point \((\mu_0,0)\) on the \(q_u\)-axis such
that \(P_\mathcal{O}(\mu_0) > \pi\), the orbit will no longer return to the
\(q_u\)-axis. 
Instead, it will escape \(\mathcal{O}\) entirely, which means it must 
escape any compact subset of \(\mathcal{O}\) in
finite time. 
We conclude that the acrobot is gaining energy on \(\mathcal{O}\).

Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
decreasing.
Hence, the orbit in \(\mathcal{O}\) spirals towards the origin, which means the
constraint is dissipating energy on \(\mathcal{O}\).

@@

\subsection*{Proving Energy Gain}
We will prove the acrobot gains energy on the set 
\[
    \Omega := \left\{ (q_u,p_u) \in \Gamma 
     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    .
\]
% Can we use the jacobian? The spectrum is hard to find.
% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
% therefore exit D and enter W.

>>>>

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
    Let 
    \[
        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
        ,
    \]
    be the energy of the simple pendulum obtained by setting \(I = 0\).
\begin{enumerate}
    \item For all \(\epsilon > 0\), there exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on the set
    \[
        \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma 
        \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0) \right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}_\epsilon\).
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right[\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \Omega := \left\{(q_u,p_u) \in \Gamma 
            \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
\end{enumerate}
\end{thm}
%\footnotetext{In fact, we can be more precise than this. 
%    Letting \(\epsilon > 0\), we need 
%    \(I \leq \min\left\{I_1(\epsilon),I_2\}\) where
%    \(I_1(\epsilon) = \frac{\tan(\epsilon)}{(\pi-\epsilon)}
%    \left(\frac{m^2gl^3(\cos(\epsilon)-2\sin(\epsilon))}{\pi-\epsilon}
%    \left(\frac{2\bar{q}_a\tan(\epsilon)}{(1+\tan(\epsilon)^2)(\pi-\epsilon)}+5\right) 
%    \right)^{-1/2}\)
%    and \(I_2 = \sqrt{\frac{3}{10 m^2gl^3\bar{q}_a^2}}\).
%    See Apendix \ref{apx:I-bounds} for details.
%}
The region \(\mathcal{O}\) is the oscillation domain of the simple pendulum
obtained by setting \(q_a = 0\);
\(\Omega\) includes \(\mathcal{O}\), alongside all rotations of this 
pendulum with momentum less than \(\bar{\mu}\). 
Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
that the acrobot will gain enough energy to begin performing giant motion;
the second result guarantees can achieve giants with momentum
\(|p_u| < \bar{\mu}\).

@@

\section{Proving the Acrobot Gains Energy}\label{sec:acrobot-proof}
% TODO: Prove this in sections
% TODO: You don't have to use the proof of theta_dot being positive. Stick to
% Manfredi's proof, and when you say "for I small enough" put in a footnote that
% says "in fact, we can be more precise and set bounds for I given as I <
% min{...}. See appendix for more detail."
% Then, we can have an appendix where we derive the bounds on I so that
% theta_dot is always positive. This will help the reader keep track and not
% bother them too much.
Proving Theorem \ref{thm:acrobot-energy-stabilization} is not a simple task.
In an effort to make the proof as clear as possible, we will break it down into
the following segments:
\begin{enumerate}
    \item Background on perturbation theory.
    \item Perturbation analysis for oscillations.
    \item Energy gain for oscillations.
    \item Perturbation analysis for rotations.
    \item Energy gain for rotations.
\end{enumerate}

@@

\citet{khalil_nonlinear} considers a system of the form

\begin{equation}\label{eqn:khalil-setup}
    \begin{cases}
        \dot{x} = f(t,x,I), \\
        x(t_0) = \eta(I),
    \end{cases}
\end{equation}
where 
\(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^n\) is smooth
\footnote{Khalil actually considers \(f(t,x,\epsilon)\) which is ``sufficiently
    smooth". We replace \(\epsilon\) with \(I\) and assume smoothness to
    more easily connect his theory to the acrobot constraint.}.
In the context of the acrobot, \(x\) is the set of coordinates which make
our analysis convenient;
\(I\) is a parameter of the system;
the function \(f\) is the constrained dynamics in \(x\)-coordinates;
and \(\eta(I) \equiv \eta_0\) is a constant initial condition.

Setting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use this to compute the solution \(x_1(t,\eta_0)\) of
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}
Finally, one can compute the first-order Taylor approximation
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\), where the remainder
term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie,
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    .
\]
We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}\label{thm:khalil-perturbation}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem ~\eqref{eqn:khalil-perturbation-nominal} has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(I,\, |I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to ~\eqref{eqn:khalil-setup} satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_0)\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
This pendulum oscillates whenever the nominal energy
~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\). 
Hence, the domain of oscillations for the nominal pendulum is given by
\[
    \mathcal{O} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(\pi,0)\right\}
    ,
\]
where \(\Gamma \simeq \SxR\) is the constraint manifold of the
acrobot.
Recall from Figure \ref{fig:pendulum-level-sets} that orbits of a pendulum 
are level sets of \(E\). 
Oscillations of a pendulum look like ellipses contained in \(\mathcal{O}\).
Since perturbation theory requires us to analyze the nominal pendulum, we will
make our analysis convenient by changing coordinates into an
``action" \(\mu > 0\) which remains constant on level sets of \(E\), 
along with an ``angle" \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
\(\mathcal{O}\).

The energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis (as in Figure \ref{fig:mu-intersection}) 
and \(q_u \in ]-\pi,\pi[\) on \(\mathcal{O}\), so we set \(\mu \in ]0,\pi[\).
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : \mathcal{O} \backslash \{(0,0)\} &\rightarrow \Sone \times \, ]0,\pi[, \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
The dynamics of the nominal pendulum arise from evaluating
\(f_\mu\) and \(f_\alpha\) at \(I = 0\).
MATLAB's symbolic toolbox evaluates the nominal dynamics as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu} &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}
            {\mu^2 \sin(\alpha)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\).
By continuity of ~\eqref{eqn:acrobot-mu-dot}, 
\(\dot{\alpha}\) remains positive on \(\mathcal{O}\) for \(I\) small enough.

@@

In the spirit of perturbation analysis, we expand the solution of
~\eqref{eqn:muhat-dot} into
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha,\mu_0,I) = \hat{\mu}_0(\alpha,\mu_0) + I
    \hat{\mu}_1(\alpha,\mu_0)
    + R(\alpha,\mu_0,I)
    ,
\end{equation}
where \(R(\alpha,\mu_0,I)\) is smooth and \(O(I^2)\).
From ~\eqref{eqn:khalil-perturbation-nominal} we know that 
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    ,
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}, we find that
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

To prove this intuitive argument rigorously, we define the Poincar\'{e} section
\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

@@

We can therefore state that, for each \(\mu_0 \in \, ]0,\pi[\), there exists
\(I\) such that
\[
    P_\mathcal{O}(\mu_0) = \mu_0\left(1+ IKQ(\mu_0)\right) + R(\pi,\mu_0,I) > 0
    .
\]
Unfortunately, Khalil's perturbation theory does not guarantee this \(I\) is the
same for all \(\mu_0\).

We deal with this issue by taking the compact set
\[
    \bar{\mathcal{O}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(q_u,p_u) \leq E(\pi,0) \right\}
    .
\]
The remainder term \(R(\pi,\mu_0,I)\) is smooth and of order \(I^2\),
which means it can be written in the form 
\(R(\pi,\mu_0,I) = I^2\tilde{R}(\mu_0,I)\)
for some \(\tilde{R}\) a smooth function.
Since \(\bar{O}\) is compact, \(\tilde{R}\) attains its minimum
\(\underbar{r}\) on \(\bar{O}\).
Hence, \(R(\pi,\mu_0,I)\) is bounded below by the parabola
\(\underbar{r}I^2\), which means
\begin{equation}\label{eqn:acrobot-Po-lowerbound}
    P_\mathcal{O}(\mu_0) \geq \mu_0\left(1 + IKQ(\mu_0)\right) + I^2\underbar{r}
    .
\end{equation}
Because \(I\) is assumed to be small, the term \(I^2\) shrinks rapidly.
If \(\underbar{r} < 0\), we can find a value of \(I\) which overcomes
the effect of \(I^2\underbar{r}\) and make \(P_\mathcal{O}\) positive.
Furthermore, ~\eqref{eqn:acrobot-Po-lowerbound} implies that, for fixed \(I\),
\(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
any \(\epsilon > 0\), there exists \(I(\epsilon)\) small enough so that
the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
\(\mu_0 \in ]\epsilon,\pi[\) and 
\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).

\subsection*{Energy gain on \(\mathcal{O}_\epsilon\)}
To simplify notation, denote by \(x(t) := (q_u(t),p_u(t))\) the orbit of the
acrobot with initial condition \(x(0)\).
Let \(\epsilon > 0\) and define
\[
    \mathcal{O}_\epsilon := \left\{(q_u,p_u) \in \Gamma
    \mid E(\epsilon,0) < E(q_u,p_u) < E(\pi,0)\right\}
    ,
\]
with \(\bar{\mathcal{O}}_\epsilon\) its (compact) closure.

Since \(\bar{\mathcal{O}}_\epsilon \subset \mathcal{O}\), we still have 
\(\dot{\alpha} > 0\), which means \(x(t)\) spirals clockwise around the origin.
If this orbit remains in \(\mathcal{O}\), it must intersect
the \(q_u\) axis at some radius \(\mu = \delta\)
which depends on the initial condition \(x(0)\).
Take 
\[
    \underbar{\delta} = 
    \min\limits_{x(0) \in \bar{\mathcal{O}}_\epsilon} \delta(x(0))
    ,
\] 
which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
From the previous perturbation analysis,
there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
is positive for all \(\mu_0 \in ]\underbar{\delta},\pi[\).

Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
intersects the \(q_u\)-axis at some point \(x(t_0) = (m_0, 0)\).
The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
\(|m_0| > \underbar{\delta}\).
As the acrobot continues its oscillation, it might exit
\(\mathcal{O}\) if it has gained enough energy;
if not, it will intersect the \(q_u\) axis again at
a new point \(x(t_1) = (m_1,0)\).
Theorem \ref{thm:khalil-perturbation} guarantees that \(|m_1|\) differs
from \(P_\mathcal{O}(|m_0|)\) only up to order \(I^2\), 
which means \(|m_1| > |m_0|\).
From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
initializing the acrobot at \(x(t_i) = (m_i, 0)\), the
next intersection with the \(q_u\) axis will be at
\(x(t_{i+1}) = (m_{i+1},0)\) with \(|m_{i+1}| > |m_i|\).

Because the Poincar\'{e} section is monotonically increasing, 
the orbit will eventually reach an intersection point \(x(t_N) = (m_N,0)\)
such that \(P_\mathcal{O}(|m_N|) > \pi\). 
At this point, the orbit will no longer return to the
\(q_u\)-axis.
Instead, it will escape \(\mathcal{O}\) entirely; 
in doing so, it also escapes \(\mathcal{O}_\epsilon\).
Hence, the orbit must escape any compact subset of \(\mathcal{O}\) in finite
time, which means the acrobot is gaining energy on \(\mathcal{O}_\epsilon\).

Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
decreasing.
Hence, the orbit in \(\mathcal{O}_\epsilon\) spirals clockwise towards the origin, 
which means the constraint is dissipating energy.

@@

\subsection*{Energy Gain on \(\Omega\)}
We will prove the acrobot gains energy on the set 
\[
    \Omega := \left\{ (q_u,p_u) \in \Gamma 
     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    .
\]
% Can we use the jacobian? The spectrum is hard to find.
% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
% therefore exit D and enter W.
